---
title: "GREEEN tallgrass prairie restoration: RLQ analyses"
author: "Markus Bauer"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    dev: "png"
    fig_width: 7
    fig_height: 5
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
library(tinytex)
knitr::opts_chunk$set(echo = TRUE)
```

Technichal University of Munich, TUM School of Life Sciences, Chair of
Restoration Ecology, Emil-Ramann-Straße 6, 85354 Freising, Germany

\*[markus1.bauer\@tum.de](mailto:markus1.bauer@tum.de)

ORCiD ID: [0000-0001-5372-4174](https://orcid.org/0000-0001-5372-4174)
<br>
[Google Scholar](https://scholar.google.de/citations?user=oHhmOkkAAAAJ&hl=de&oi=ao)
<br>
GitHub: [markus1bauer](https://github.com/markus1bauer)
<br>
GitHub: [TUM-Restoration-Ecology](https://github.com/TUM-Restoration-Ecology)

# Preparation

#### Packages

```{r libraries, message = FALSE}
library(here)
library(tidyverse)
library(ade4)
```

```{r echo = FALSE}
rm(list = ls())
```

#### Load data

```{r load-data}
sites <- read_csv(
  here("data", "processed", "data_processed_sites.csv"),
  col_names = TRUE, na = c("na", "NA", ""), col_types =
    cols(
    .default = "?"
  )) %>% 
  arrange(id_plot_year) %>%
  filter(
    # year != "2015",
    site == "Lux Arbor",
    richness_type == "seeded_richness",
    !(treatment_id %in% c("2", "4"))
  ) %>%
  select(
    id_plot_year, year, seeded_pool, water_cap, cover_total,
    cover_non_seeded
    )

traits <- read_csv(
  here("data", "processed", "data_processed_traits.csv"),
  col_names = TRUE, na = c("na", "NA", ""), col_types = cols(.default = "?")
  ) %>%
  filter(seeded == "1") %>%
  select(accepted_name, sla, height, seedmass) %>%
  drop_na()

species <- read_csv(
  here("data", "processed", "data_processed_species.csv"),
  col_names = TRUE, na = c("na", "NA", ""), col_types = cols(.default = "?")
  ) %>%
  filter(plot_size == "1") %>%
  select(id_plot_year, accepted_name, abundance) %>%
  semi_join(sites, by = "id_plot_year") %>%
  semi_join(traits, by = "accepted_name") %>%
  arrange(id_plot_year)

species %>%
  group_by(id_plot_year) %>%
  count()

traits %>% anti_join(species, by = "accepted_name")
  
traits <- traits %>%
  semi_join(species, by = "accepted_name") %>%
  column_to_rownames(var = "accepted_name")

sites <- sites %>%
  semi_join(species, by = "id_plot_year") %>%
  column_to_rownames(var = "id_plot_year")

species <- species %>%
  pivot_wider(
    names_from = "accepted_name", values_from = "abundance", values_fill = 0
    ) %>%
  column_to_rownames(var = "id_plot_year")
```

# Statistics

## Descriptive analyses

```{r}
sites %>%
  group_by(year) %>%
  count()
```


## Ordinations

```{r warning=FALSE}
dudi_l <- dudi.coa(species, scannf = FALSE)
dudi_r <- dudi.hillsmith(
  sites, 
  row.w = dudi_l$lw,
  scannf = FALSE
  )
dudi_q <- dudi.pca(
  traits, 
  row.w = dudi_l$cw,
  scannf = FALSE
  )

```


## RLQ analyses

```{r load-models, collapse = TRUE}
m1 <- rlq(
  dudiR = dudi_r, 
  dudiL = dudi_l, 
  dudiQ = dudi_q,
  scannf = FALSE
  )
```

## Global test

```{r}
randtest(m1, nrepet = 999, modeltype = 6)
```

## Traits by environment crossed table

```{r}
m1$tab 
```

# Plots

```{r}
s.label(m1$lR)
s.label(m1$lQ)
s.arrow(m1$l1)
s.arrow(m1$c1)
```

# Save

```{r}
scores_sites <- m1$lR %>%
  mutate(type = "sites") %>%
  rename(axis1 = 1, axis2 = 2)
scores_species <- m1$lQ %>%
  mutate(type = "species") %>%
  rename(axis1 = 1, axis2 = 2)
scores_environment <- m1$l1 %>%
  mutate(type = "environment") %>%
  rename(axis1 = 1, axis2 = 2)
scores_traits <- m1$c1 %>%
  mutate(type = "traits") %>%
  rename(axis1 = 1, axis2 = 2)

scores <- bind_rows(
  scores_sites, scores_species, scores_environment, scores_traits
  ) %>%
  rownames_to_column(var = "id")

write_csv(
  scores, here("data", "processed", "data_processed_scores.csv")
)
```


# Session info

```{r session-info, echo = FALSE}
sessionInfo()
```
