---
title: "GREEEN tallgrass prairie restoration: Seeded richness ~ Species pool size (NW Station)"
author: "MB"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    dev: "png"
    fig_width: 7
    fig_height: 5
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
  )
```

<br/>
<br/>

> **NOTE:** To compare different models, you only have to change the models in the section 'Load models'

# Preparation

Protocol of data exploration (Steps 1-8) used from Zuur et al. (2010) Methods Ecol Evol [DOI: 10.1111/2041-210X.12577](https://doi.org/10.1111/2041-210X.12577)

#### Packages

```{r libraries, message = FALSE}
library(here)
library(tidyverse)
library(ggbeeswarm)
library(patchwork)
library(lme4)
library(glmmTMB)
library(DHARMa)
library(emmeans)
```

```{r echo = FALSE}
rm(list = ls())
```

#### Load data

```{r load-data}
sites <- read_csv(
  here("data", "processed", "data_processed_sites.csv"),
  col_names = TRUE, na = c("na", "NA", ""), col_types = cols(
    .default = "?",
    id_plot = "f",
    year = "f",
    seeding_time = col_factor(
      levels = c("unseeded", "fall", "spring"), ordered = FALSE
      ),
    seeded_pool = col_factor(
      levels = c("0", "6", "12", "18", "33"), ordered = TRUE
      ),
    treatment_id = "f",
    richness_type = "f"
  )
) %>%
  filter(
    site == "NW Station",
    richness_type == "seeded_richness",
    !(treatment_id %in% c("2", "4")),
    !(year == 2016)
  ) %>%
  mutate(y = richness_1qm + richness_25qm)
```

# Statistics

## Data exploration

### Means and deviations

```{r means}
Rmisc::CI(sites$y, ci = .95)
median(sites$y)
sd(sites$y)
quantile(sites$y, probs = c(0.05, 0.95), na.rm = TRUE)
sites %>% count(site, year)
sites %>% count(seeded_pool)
sites %>% count(seeding_time)
sites %>% count(herbicide)
```

### Graphs of raw data (Step 2, 6, 7)

```{r data-exploration, echo = FALSE, warning = FALSE, message = FALSE}
ggplot(sites, aes(y = y, x = year)) +
  geom_quasirandom(color = "grey") +
  geom_boxplot(fill = "transparent") +
  facet_grid(~ site) +
  labs(y = "Seeded species richness (25qm)", x = "Survey year")

ggplot(
  data = sites,
  aes(y = y, x = seeded_pool, fill = seeding_time) # seeding time = herbicide treatment
) +
  geom_quasirandom(aes(color = seeding_time), dodge.width = .8, alpha = .8) +
  geom_boxplot(alpha = .5) +
  facet_grid(~ site) +
  labs(y = "Seeded species richness (25qm)", x = "Species pool [#]")

ggplot(
  data = sites, # herbicide = seeding_time
  aes(y = y, x = seeded_pool, fill = seeding_time, color = seeding_time)
) +
  geom_quasirandom(dodge.width = .8, alpha = .8) +
  geom_boxplot(fill = "transparent") +
  facet_grid(site ~ year) +
  labs(y = "Seeded species richness (25qm)", x = "Species pool [#]")
```

### Outliers, zero-inflation, transformations? (Step 1, 3, 4)

```{r outliers, echo = FALSE, warning = FALSE}
plot1 <- ggplot(sites, aes(x = site, y = y)) + geom_quasirandom()
plot2 <- ggplot(sites, aes(x = y)) + geom_histogram(binwidth = 0.7)
plot3 <- ggplot(sites, aes(x = y)) + geom_density()
plot4 <- ggplot(sites, aes(x = log(y))) + geom_density()
(plot1 + plot2) / (plot3 + plot4)
```

### Check collinearity part 1 (Step 5)

Exclude r > 0.7
<br>
Dormann et al. 2013 Ecography [DOI: 10.1111/j.1600-0587.2012.07348.x](https://doi.org/10.1111/j.1600-0587.2012.07348.x)

```{r collinearity, warning = FALSE, message = FALSE}
# sites %>%
#   select(where(is.numeric), -y, -starts_with("cwm.")) %>%
#   GGally::ggpairs(
#     lower = list(continuous = "smooth_loess")
#     ) +
#   theme(strip.text = element_text(size = 7))

# -> no continuous variables
```

## Models

> **NOTE:** Only here you have to modify the script to compare other models

```{r load-models, collapse = TRUE}
load(file = here("outputs", "models", "model_richness_pool_nws_1.Rdata"))
m_1 <- m1
# m_2 <- m2
```

```{r formulas, collapse = TRUE}
formula(m_1)
class(m_1)
# formula(m_2)
# class(m_2)
# m_1@call
# m_2@call
```

## Model check

### DHARMa

```{r dharma_all, collapse = TRUE, warning = FALSE}
simulation_output_1 <- simulateResiduals(m_1, plot = TRUE)
# simulation_output_2 <- simulateResiduals(m_2, plot = TRUE)
```

```{r dharma_single, collapse = TRUE, fig.height = 3, fig.width = 4.5}
plotResiduals(
  simulation_output_1$scaledResiduals,
  factor(sites %>% filter(!(is.na(water_cap))) %>%  pull(seeded_pool))
  )
# plotResiduals(simulation_output_2$scaledResiduals, factor(sites$seeded_pool))
plotResiduals(
  simulation_output_1$scaledResiduals,
  factor(sites %>% filter(!(is.na(water_cap))) %>%  pull(herbicide))
  )
# plotResiduals(simulation_output_2$scaledResiduals, sites$herbicide)
plotResiduals(
  simulation_output_1$scaledResiduals,
  factor(sites %>% filter(!(is.na(water_cap))) %>%  pull(seeding_time))
  )
# plotResiduals(simulation_output_2$scaledResiduals, sites$seeding_time)
plotResiduals(
  simulation_output_1$scaledResiduals,
  factor(sites %>% filter(!(is.na(water_cap))) %>%  pull(year))
  )
# plotResiduals(simulation_output_2$scaledResiduals, sites$year)
plotResiduals(
  simulation_output_1$scaledResiduals,
  factor(sites %>% filter(!(is.na(water_cap))) %>%  pull(water_cap))
  )
# plotResiduals(simulation_output_2$scaledResiduals, sites$water_cap)
```
#### Test for overdispersion

```{r overdispersion, collapse = TRUE, fig.height = 3, fig.width = 4.5}
testDispersion(simulation_output_1)
# testDispersion(simulation_output_2)
```

### Check collinearity part 2 (Step 5)

Remove VIF > 3 or > 10
<br>
Zuur et al. 2010 Methods Ecol Evol [DOI: 10.1111/j.2041-210X.2009.00001.x](https://doi.org/10.1111/j.2041-210X.2009.00001.x)

```{r vif}
car::vif(m_1)
# car::vif(m_2)
```

## Model comparison

### <i>R</i><sup>2</sup> values

```{r r2, collapse = TRUE}
MuMIn::r.squaredGLMM(m_1)
# MuMIn::r.squaredGLMM(m_2)
```

### AICc

Use AICc and not AIC since ratio n/K < 40
<br>
Burnahm & Anderson 2002 p. 66
ISBN: 978-0-387-95364-9

```{r aicc, collapse = TRUE}
# MuMIn::AICc(m_1, m_2) %>%
#   arrange(AICc)
```

## Predicted values

### Summary table

```{r summary}
car::Anova(m_1, type = 3)
summary(m_1)
```

### Forest plot

```{r predicted_values, fig.height = 6}
dotwhisker::dwplot(
  list(m_1),
  ci = 0.95,
  show_intercept = FALSE,
  vline = geom_vline(xintercept = 0, colour = "grey60", linetype = 2)) +
  theme_classic()
```

### Effect sizes

Effect sizes of chosen model just to get exact values of means etc. if
necessary.

```{r effect-sizes-1, message = FALSE}
(emm <- emmeans(
  m_1,
  revpairwise ~ seeding_time | seeded_pool,
  type = "response"
  ))
plot(emm, comparison = TRUE)
```

# Session info

```{r session-info, echo = FALSE}
sessionInfo()
```
