---
title: "Garchinger Heide and restoration sites: <br> NMDS and PERMANOVA"
author: "<b>Markus Bauer</b> <br>"
date: "<b>`r format(Sys.time(), '%Y-%m-%d')`</b>"
output:
  github_document:
    toc: true
    toc_depth: 3
    dev: png
    fig_width: 7
    fig_height: 5
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
  )
```

Technichal University of Munich, TUM School of Life Sciences, Chair of
Restoration Ecology, Emil-Ramann-Straße 6, 85354 Freising, Germany

\*[markus1.bauer\@tum.de](mailto:markus1.bauer@tum.de)

ORCiD ID: [0000-0001-5372-4174](https://orcid.org/0000-0001-5372-4174)
<br>
[Google Scholar](https://scholar.google.de/citations?user=oHhmOkkAAAAJ&hl=de&oi=ao)
<br>
GitHub: [markus1bauer](https://github.com/markus1bauer)
<br>
GitHub: [TUM-Restoration-Ecology](https://github.com/TUM-Restoration-Ecology)

# Preparation

#### Packages

```{r libraries, message = FALSE}
library(here)
library(tidyverse)
library(vegan)
```

```{r echo = FALSE}
rm(list = ls())
```

#### Load data

```{r load-data}
sites <- read_csv(
  here("data", "processed", "data_processed_sites.csv"),
  col_names = TRUE, na = c("na", "NA", ""), col_types =
    cols(
    .default = "?"
  )) %>% 
  arrange(id)

species <- read_csv(
  here("data", "processed", "data_processed_species.csv"),
  col_names = TRUE, na = c("na", "NA", ""), col_types = cols(.default = "?")
  ) %>%
  pivot_longer(-accepted_name, names_to = "id", values_to = "value") %>%
  semi_join(sites, by = "id") %>%
  arrange(id) %>%
  pivot_wider(names_from = "accepted_name", values_from = "value") %>%
  column_to_rownames(var = "id")
```

# Statistics

## Descriptive statistics

```{r warning=FALSE}
sites %>%
  ggplot(aes(x = as.numeric(year_hay_transfer), fill = mowing_date)) +
  geom_histogram()
sites %>%
  filter(!(is.na(year_hay_transfer) | year_hay_transfer == "no")) %>%
  ggplot(aes(x = mowing_date_start, fill = mowing_date)) +
  geom_histogram()
sites %>%
  group_by(treatment) %>%
  count(esy) %>%
  group_by(treatment) %>%
  mutate(ratio = round(n / sum(n), digits = 2))
```


## Models

```{r load-models, collapse = TRUE}
# set.seed(9)
# ordi <- metaMDS(
#   species, dist = "bray", k = 2, engine = "monoMDS", binary = FALSE,
#   try = 99, previous.best = TRUE, na.rm = TRUE
#   )
# save(ordi, file = here("outputs", "models", "model_nmds.Rdata"))
base::load(here("outputs", "models", "model_nmds.Rdata"))
```

```{r formulas, collapse = TRUE}
ordi
```

## Model check

```{r model-check, collapse = TRUE, warning = FALSE}
stressplot(ordi)
goodness_of_fit <- goodness(ordi)
plot(ordi, type = "t", main = "Goodness of fit")
points(ordi, display = "sites", cex = goodness_of_fit * 300)
```


## Vectors

```{r r2, collapse = TRUE}
ef_vector1 <- envfit(
  ordi ~ height_vegetation + cover_vegetation + grass_cover + graminoid_cover,
  data = sites,
  permu = 999,
  na.rm = TRUE
)
ef_vector1
plot(ordi, type = "n")
plot(ef_vector1, add = TRUE, p. = .99)
# save(ef_vector1, file = here("outputs", "models", "model_nmds_envfit_vector1.Rdata"))
```

## Factors

```{r aicc, collapse = TRUE}
ef_factor1 <- envfit(
  ordi ~  treatment,
  data = sites, permu = 999, na.rm = TRUE
)
ef_factor1
```

# PERMDISP and PERMANOVA

## PERMDISP

```{r permdisp}
dispersion <- betadisper(dist(species), sites$treatment)
permutest(dispersion)
boxplot(dispersion)
TukeyHSD(dispersion)
plot(dispersion, hull = FALSE, ellipse = TRUE, label = TRUE)
```

## PERMANOVA

Exclude reference sites (control) from 2003 because of high dispersion within this group

```{r permanova}
data <- species %>%
  rownames_to_column(var = "id") %>%
  filter(!(str_detect(id, "2003roeder")))
sites2 <- sites %>%
  semi_join(data, by = "id")
species2 <- data %>%
  column_to_rownames(var = "id")

(permanova <- adonis2(
  species2 ~ treatment, data = sites2,
  permutations = 999,
  method = "bray")
)
```

# Session info

```{r session-info, echo = FALSE}
sessionInfo()
```
