---
title: "Analysis of Bauer et al. (2023) bioRxiv: <br> Favourable Conservation Status (FCS)"
author: "<b>Markus Bauer</b> <br>"
date: "<b>`r format(Sys.time(), '%Y-%m-%d')`</b>"
output:
  github_document:
    toc: true
    toc_depth: 3
    dev: png
    fig_width: 7
    fig_height: 5
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
  )
```

<br/> <br/> <b>Markus Bauer</b>

Technichal University of Munich, TUM School of Life Sciences, Chair of
Restoration Ecology, Emil-Ramann-Straße 6, 85354 Freising, Germany

[markus1.bauer\@tum.de](mailto:markus1.bauer@tum.de)

ORCiD ID: [0000-0001-5372-4174](https://orcid.org/0000-0001-5372-4174)
<br> [Google
Scholar](https://scholar.google.de/citations?user=oHhmOkkAAAAJ&hl=de&oi=ao)
<br> GitHub: [markus1bauer](https://github.com/markus1bauer)

To compare different models, you only have to change the models in
section `Load models`

# Preparation

Favourable Conservation Status (FCS) sensu Helm et al. (2015) Divers
Distrib [DOI: 10.1111/ddi.12285](https://doi.org/10.1111/ddi.12285)

Bayesian analysis motivated by Applestein et al. (2021) Restor Ecol
[DOI: 10.1111/rec.13596](https://doi.org/10.1111/rec.13596)

Analysis guided by <br> <b>BARG</b> (Bayesian Analysis Reporting
Guidelines): Kruschke (2021) Nat Hum Behav [DOI:
10.1038/s41562-021-01177-7](https://doi.org/10.1038/s41562-021-01177-7)
<br> Model check: Gabry et al. (2019) J R Stat Soc A Stat [DOI:
10.1111/rssa.12378](https://doi.org/10.1111/rssa.12378) <br> Priors:
Lemoine (2019) Oikos [DOI:
10.1111/oik.05985](https://doi.org/10.1111/oik.05985) <br> Model check:
Deapoli & Schoot (2017) Psychol Methods [DOI:
10.1037/met0000065](https://doi.org/10.1037/met0000065)

#### Packages (BARG 2.A)

```{r libraries, message = FALSE}
library(here)
library(tidyverse)
library(ggbeeswarm)
library(patchwork)
library(brms)
#remotes::install_github("Pakillo/DHARMa.helpers")
library(DHARMa.helpers)
library(bayesplot)
library(loo)
library(emmeans)
```

```{r echo = FALSE}
rm(list = ls())
theme_mb <- function() {
  theme(
    panel.background = element_rect(fill = "white"),
    text = element_text(size = 9, color = "black"),
    strip.text = element_text(size = 10),
    axis.text = element_text(angle = 0, hjust = 0.5, size = 9,
                             color = "black"),
    axis.title = element_text(angle = 0, hjust = 0.5, size = 9,
                              color = "black"),
    axis.line = element_line(),
    legend.key = element_rect(fill = "white"),
    legend.text = element_text(size = 9),
    legend.position = "right",
    legend.margin = margin(0, 0, 0, 0, "cm"),
    plot.margin = margin(0, 0, 0, 0, "cm")
  )
}
```

#### Load data

```{r load-data}
sites <- read_csv(
  here("data", "processed", "data_processed_sites.csv"),
  col_names = TRUE, na = c("na", "NA", ""), col_types = cols(
    .default = "?",
    id_plot_year = "f",
    id_plot = "f",
    site = col_factor(
      levels = c("NW Station", "Lux Arbor", "SW Station"), ordered = FALSE
    ),
    year = "f",
    seeding_time = col_factor(
      levels = c("unseeded", "fall", "spring"), ordered = FALSE
      ),
    herbicide = col_factor(levels = c("0", "1"), ordered = FALSE),
    seeded_pool = col_factor(
      levels = c("0", "6", "12", "18", "33"), ordered = TRUE
      ),
    treatment_id = "f",
    treatment_description = "c",
    richness_type = "f"
  )
) %>%
  filter(
    year %in% c("2015", "2016", "2017", "2018"),
    richness_type == "seeded_richness",
    !(treatment_id %in% c("2", "4"))
  ) %>%
  select(
    id_plot_year, id_plot, site, year, herbicide, seeding_time, seeded_pool,
    richness_1qm, richness_25qm, treatment_id
    ) %>%
  mutate(y = richness_1qm + richness_25qm)
sites
```

# Statistics

## Data exploration

### Means and deviations

```{r means}
Rmisc::CI(sites$y, ci = .95)
median(sites$y)
sd(sites$y)
quantile(sites$y, probs = c(0.05, 0.95), na.rm = TRUE)
sites %>% count(site, year)
sites %>% count(seeded_pool)
sites %>% count(seeding_time)
sites %>% count(herbicide)
```

### Graphs of raw data

```{r data-exploration, echo = FALSE}
ggplot(sites, aes(y = y, x = year)) +
  geom_quasirandom(color = "grey") +
  geom_boxplot(fill = "transparent") +
  facet_grid(~ site) +
  labs(y = "Seeded species richness (25qm)", x = "Survey year") +
  theme_mb()
ggplot(
  sites, aes(y = y, x = year, fill = seeding_time) # herbicide = seeding_time
  ) +
  geom_quasirandom(aes(color = seeding_time), dodge.width = .8, alpha = .8) +
  geom_boxplot(alpha = .3) +
  facet_grid(~ site) +
  labs(y = "Seeded species richness (25qm)", x = "Herbicide treatment") +
  theme_mb()
ggplot(
  sites, aes(y = y, x = year, fill = seeded_pool)
  ) +
  geom_quasirandom(aes(color = seeded_pool), dodge.width = .8, alpha = .8) +
  geom_boxplot(alpha = .3) +
  facet_grid(~ site) +
  labs(y = "Seeded species richness (25qm)", x = "Species pool [#]") +
  theme_mb()
ggplot(
  data = sites,
  aes(y = y, x = seeded_pool, fill = seeding_time)
  ) +
  geom_quasirandom(aes(color = seeding_time), dodge.width = .8, alpha = .8) +
  geom_boxplot(alpha = .3) +
  facet_grid(~ year) +
  labs(y = "Seeded species richness (25qm)", x = "Species pool [#]") +
  theme_mb()
ggplot(
  data = sites,
  aes(y = y, x = seeded_pool, fill = seeding_time)
  ) +
  geom_quasirandom(aes(color = seeding_time), dodge.width = .8, alpha = .8) +
  geom_boxplot(alpha = .5) +
  facet_grid(~ site) +
  labs(y = "Seeded species richness (25qm)", x = "Species pool [#]") +
  theme_mb()
ggplot(
  data = sites,
  aes(y = y, x = seeded_pool, fill = seeding_time)
) +
  geom_quasirandom(aes(color = seeding_time), dodge.width = .8, alpha = .8) +
  geom_boxplot(alpha = .5) +
  facet_grid(site ~ year) +
  labs(y = "Seeded species richness (25qm)", x = "Species pool [#]") +
  theme_mb()
```

### Outliers, zero-inflation, transformations?

```{r outliers, echo = FALSE}
plot1 <- ggplot(sites, aes(x = site, y = y)) + geom_quasirandom()
plot2 <- ggplot(sites, aes(x = y)) + geom_histogram(binwidth = 0.7)
plot3 <- ggplot(sites, aes(x = y)) + geom_density()
plot4 <- ggplot(sites, aes(x = log(y))) + geom_density()
(plot1 + plot2) / (plot3 + plot4)
```

## Models

### Load models (BARG 6.G)

Only here you have to modify the script to compare other models

```{r load-models, collapse = TRUE}
load(file = here("outputs", "models", "model_pool_1_flat.Rdata"))
load(file = here("outputs", "models", "model_pool_full.Rdata"))
# BARG 5.A/B/C
load(file = here("outputs", "models", "model_pool_1_flat.Rdata"))
load(file = here("outputs", "models", "model_pool_1_prior.Rdata"))
m_1 <- m1_flat
m_2 <- m_full
m_prior <- m1_prior
m_flat <- m1_flat
```

### Model specifications (BARG 1.D)

```{r formulas, collapse = TRUE}
m_1$formula
m_2$formula
```

```{r families, collapse = TRUE}
m_1$family
m_2$family
```

Markov Chain Monte Carlo (MCMC) method used with No-U-Turn Sampler
(NUTS) which is an extension of Hamiltonian Monte Carlo (HMC).

Amount of chains for MCMC

```{r model-specification-chains, collapse = TRUE}
m_1$fit@sim$chains
m_2$fit@sim$chains
```

Total amount of iterations for MCMC

```{r model-specification-iterations, collapse = TRUE}
m_1$fit@sim$iter
m_2$fit@sim$iter
```

Amount of iterations before burn-in

```{r model-specification-warmup, collapse = TRUE}
m_1$fit@sim$warmup
m_2$fit@sim$warmup
```

Thinning rate

```{r model-specification-thinning, collapse = TRUE}
m_1$fit@sim$thin
m_2$fit@sim$thin
```

### Preparation for analysis

```{r analysis-preparation}
# Chose variables
variables <- c(
      "Intercept",
      "b_herbicide1",
      "b_seeded_pool.C",
      "b_seeded_pool.L",
      "b_seeded_pool.Q",
      "b_siteLuxArbor",
      "b_siteSWStation",
      "b_seeded_pool.L:herbicide1:siteLuxArbor"#,
      #"sd_year__Intercept"#,
      #"sigma"
    )
# Subset draws
posterior1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(variable = variables)
posterior2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::subset_draws(variable = variables)
posterior_prior <- m_prior %>%
  posterior::as_draws() %>%
  posterior::subset_draws(variable = variables)
posterior_flat <- m_flat %>%
  posterior::as_draws() %>%
  posterior::subset_draws(variable = variables)
# Long format of draws
hmc_diagnostics1 <- m_1 %>% brms::nuts_params()
hmc_diagnostics2 <- m_2 %>% brms::nuts_params()
y <- sites$y
# Posterior predictive distribution
yrep1 <- m_1 %>% brms::posterior_predict(draws = 500)
yrep2 <- m_2 %>% brms::posterior_predict(draws = 500)
yrep_prior <- m_prior %>% brms::posterior_predict(draws = 500)
# Summary statistics
draws1 <- m_1 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
draws2 <- m_2 %>%
  posterior::as_draws() %>%
  posterior::summarize_draws() %>%
  filter(str_starts(variable, "b_"))
```

### Priors (BARG 1.D/E)

#### Possible prior distributions

```{r possible-priors}
brms::get_prior(
  y ~ seeded_pool * herbicide * site + (1 | year), 
  data = sites
  )
data <- data.frame(x = c(-10, 10))
ggplot(data, aes(x = x)) +
  stat_function(fun = dnorm, n = 101, args = list(mean = 0, sd = 10)) +
  expand_limits(y = 0) +
  ggtitle("Normal distribution for Intercept")
ggplot(data, aes(x = x)) +
  stat_function(fun = dnorm, n = 101, args = list(mean = 0.5, sd = 10)) +
  expand_limits(y = 0) +
  ggtitle("Normal distribution for treatments")
ggplot(data, aes(x = x)) +
  stat_function(fun = dcauchy, n = 101, args = list(location = 0, scale = 1)) +
  expand_limits(y = 0) +
  ggtitle("Cauchy distribution")
ggplot(data, aes(x = x)) +
  stat_function(fun = dstudent_t, args = list(df = 3, mu = 0, sigma = 2.5)) +
  expand_limits(y = 0) +
  ggtitle(expression(Student ~ italic(t) * "-distribution"))
```

#### Prior summary (BARG 1.D)

```{r prior-summary}
m_1 %>%
  brms::prior_summary(all = FALSE) %>%
  select(prior, class, coef, group, source)
```

#### Prior predictive check (BARG 1.E)

The blue line (observed value) should be in the center of the
distribution of 500 simulated data sets.

```{r prior-predictive-check, fig.height = 3}
bayesplot::ppc_stat(y, yrep_prior[1:500, ], binwidth = 0.5) +
  coord_cartesian(xlim = c(-4, 4)) + bayesplot::theme_default()
ppc_stat_grouped(
  y, yrep_prior[1:500, ], group = sites$year, binwidth = 0.5
  ) +
  coord_cartesian(xlim = c(-4, 4)) + bayesplot::theme_default()
ppc_stat_grouped(
  y, yrep_prior[1:500, ], group = sites$site, binwidth = 0.5
  ) +
  coord_cartesian(xlim = c(-4, 4)) + bayesplot::theme_default()
ppc_stat_grouped(
  y, yrep_prior[1:500, ], group = sites$herbicide, binwidth = 0.5
  ) +
  coord_cartesian(xlim = c(-4, 4)) + bayesplot::theme_default()
ppc_stat_grouped(
  y, yrep_prior[1:500, ], group = sites$seeded_pool, binwidth = 0.5
  ) +
  coord_cartesian(xlim = c(-4, 4)) + bayesplot::theme_default()
```

## Model check

### Check computation (MCMC diagnostics, BARG 2.B/C)

#### Trace plots

The traces of the four chains should overlap each other.

```{r mcmc-trace, fig.height = 14}
bayesplot::mcmc_trace(
  posterior1, np = hmc_diagnostics1, facet_args = list(ncol = 2)
  ) + bayesplot::theme_default()
bayesplot::mcmc_trace(
  posterior2, np = hmc_diagnostics2, facet_args = list(ncol = 2)
  ) + bayesplot::theme_default()
```

#### Sampling efficency: R-hat (BARG 2.B)

R-hat values should be lower than 1.1.

```{r rhat, collapse = TRUE}
m_1 %>% brms::rhat() %>% bayesplot::mcmc_rhat() + bayesplot::theme_default() + theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
m_2 %>% brms::rhat() %>% bayesplot::mcmc_rhat() + bayesplot::theme_default() + theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
```

#### Sampling effectiveness: Effective sampling size (ESS) (BARG 2.C)

ESS should be greater than 0.1.

```{r ess, collapse = TRUE}
m_1 %>% brms::neff_ratio() %>% bayesplot::mcmc_neff() + bayesplot::theme_default() + theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
m_2 %>% brms::neff_ratio() %>% bayesplot::mcmc_neff() + bayesplot::theme_default() + theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
```

#### Pairs plot

If the technical problem of divergent transitions occurs, points would
be highlighted in green.

```{r mcmc-pairs, collapse = TRUE, fig.height = 6.5}
m_1 %>% bayesplot::mcmc_pairs(
  off_diag_args = list(size = 1.2),
  pars = c(
    "b_herbicide1", "b_seeded_pool.C", "b_seeded_pool.L",
    "b_seeded_pool.Q", "b_siteLuxArbor",
    "b_siteSWStation"
           )
)
m_2 %>% bayesplot::mcmc_pairs(
  off_diag_args = list(size = 1.2),
  pars = c(
    "b_herbicide1", "b_seeded_pool.C", "b_seeded_pool.L",
    "b_seeded_pool.Q", "b_siteLuxArbor",
    "b_seeded_pool.Q:herbicide1:siteLuxArbor",
    "b_siteSWStation", "sigma"
           )
)
```

#### Parallel coordinate plot

If the technical problem of divergent transitions occurs, points would
be highlighted in green.

```{r mcmc-parcoord, collapse = TRUE}
posterior1 %>% bayesplot::mcmc_parcoord(np = hmc_diagnostics1) + bayesplot::theme_default() + theme(axis.text.x = element_text(angle = 45))
posterior2 %>% bayesplot::mcmc_parcoord(np = hmc_diagnostics2) + bayesplot::theme_default() + theme(axis.text.x = element_text(angle = 45))
```

#### Autocorrelation check

Autocorrelation should go down to zero very fast.

```{r autocorrelation}
posterior1 %>% bayesplot::mcmc_acf(lags = 10) + bayesplot::theme_default()
posterior2 %>% bayesplot::mcmc_acf(lags = 10) + bayesplot::theme_default()
```

### Posterior predictive check (BARG 3.A)

#### Kernel density

The kernel density estimate of the observed data (dark line) compared
with estimates for 50 simulated data sets drawn from the posterior
predictive distribution (thinner and lighter lines). The dark line
should be near the simulated curves.

```{r kernel-density}
p1 <- bayesplot::ppc_dens_overlay(y, yrep1[1:50, ]) + bayesplot::theme_default()
p2 <- bayesplot::ppc_dens_overlay(y, yrep2[1:50, ]) + bayesplot::theme_default()
p1 / p2
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$site) + bayesplot::theme_default()
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$site) + bayesplot::theme_default()
p1 / p2
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$herbicide) + bayesplot::theme_default()
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$herbicide) + bayesplot::theme_default()
p1 / p2
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$seeded_pool) + bayesplot::theme_default()
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$seeded_pool) + bayesplot::theme_default()
p1 / p2
p1 <- ppc_dens_overlay_grouped(y, yrep1[1:50, ], group = sites$year) + bayesplot::theme_default()
p2 <- ppc_dens_overlay_grouped(y, yrep2[1:50, ], group = sites$year) + bayesplot::theme_default()
p1 / p2
```

#### Histograms of statistics skew

The dark line (computed from the observed data) should be in the center
of the posterior predictive distribution.

```{r histogram}
p1 <- bayesplot::ppc_stat(y, yrep1, binwidth = 0.001) + bayesplot::theme_default()
p2 <- bayesplot::ppc_stat(y, yrep2, binwidth = 0.001) + bayesplot::theme_default()
p1 / p2
p1 <- ppc_stat_grouped(y, yrep1, group = sites$site, binwidth = 0.001) + bayesplot::theme_default()
p2 <- ppc_stat_grouped(y, yrep2, group = sites$site, binwidth = 0.001) + bayesplot::theme_default()
p1 / p2
p1 <- ppc_stat_grouped(y, yrep1, group = sites$herbicide, binwidth = 0.001) + bayesplot::theme_default()
p2 <- ppc_stat_grouped(y, yrep2, group = sites$herbicide, binwidth = 0.001) + bayesplot::theme_default()
p1 / p2
p1 <- ppc_stat_grouped(y, yrep1, group = sites$seeded_pool, binwidth = 0.001) + bayesplot::theme_default()
p2 <- ppc_stat_grouped(y, yrep2, group = sites$seeded_pool, binwidth = 0.001) + bayesplot::theme_default()
p1 / p2
```

#### LOO cross-validation (Leave one out)

All Pareto-k estimates should be \< 0.7.

```{r loo}
# loo1 <- m_1 %>% brms::loo(save_psis = TRUE, moment_match = TRUE)
# loo2 <- m_2 %>% brms::loo(save_psis = TRUE, moment_match = TRUE)
# loo1
# loo2
# plot(loo1)
# plot(loo2)
```

Density of the computed leave one out probability integral transform
(LOO PIT; dark curve) compared with the posterior predictive
distribution (light curves). The LOO PIT curve should be within the
range of light curves.

```{r loo-pit, message = FALSE}
# p1 <- bayesplot::ppc_loo_pit_overlay(y, yrep1, lw = weights(loo1$psis_object)) + bayesplot::theme_default()
# p2 <- bayesplot::ppc_loo_pit_overlay(y, yrep2, lw = weights(loo2$psis_object)) + bayesplot::theme_default()
# p1 / p2
```

### DHARMa

```{r dharma, collapse = TRUE, fig.height = 5}
m_1 %>% DHARMa.helpers::dh_check_brms(integer = TRUE)
m_2 %>% DHARMa.helpers::dh_check_brms(integer = TRUE)
```

## Model comparison

### Conditional <i>R</i><sup>2</sup> values

```{r conditional-r2, collapse = TRUE}
m_1 %>% brms::bayes_R2(
  probs = c(0.05, 0.5, 0.95), re_formula =  ~ (1 | year)
  )
m_2 %>% brms::bayes_R2(
  probs = c(0.05, 0.5, 0.95), re_formula =  ~ (1 | year)
  )
```

### Marginal <i>R</i><sup>2</sup> values

```{r marginal-r2, collapse = TRUE}
m_1 %>% brms::bayes_R2(
  probs = c(0.05, 0.5, 0.95), re_formula = 1 ~ 1
  )
m_2 %>% brms::bayes_R2(
  probs = c(0.05, 0.5, 0.95), re_formula = 1 ~ 1
  )
```

### Bayes factor (BARG 3.C)

```{r bayes-factor-1, results = FALSE}
# bayes_factor <- brms::bayes_factor(m_1, m_2)
```

```{r bayes-factor-2}
# bayes_factor
```

## Posterior distributions (BARG 3.B)

### Forest plot of effect sizes (BARG 3.B/5.B)

```{r posteriors, fig.height = 7}
combined_models <- bind_rows(
  bayesplot::mcmc_intervals_data(
    posterior1, prob = 0.66, prob_outer = 0.95
    ) %>%
    mutate(model = "m_1"),
  bayesplot::mcmc_intervals_data(
    posterior2, prob = 0.66, prob_outer = 0.95
    ) %>%
    mutate(model = "m_2"),
  bayesplot::mcmc_intervals_data(
    posterior_flat, prob = 0.66, prob_outer = 0.95
    ) %>%
    mutate(model = "m_flat"),
  bayesplot::mcmc_intervals_data(
    posterior_prior, prob = 0.66, prob_outer = 0.95
    ) %>%
    mutate(model = "m_prior")
  )

pos <- position_nudge(
  y = if_else(
    combined_models$model == "m_2", -.2, if_else(
      combined_models$model == "m_flat", -.4, if_else(
        combined_models$model == "m_prior", -.6, 0
        )
      )
    )
  )

ggplot(
  data = combined_models,
  aes(x = m, y = forcats::fct_rev(factor(parameter)), color = model)
  ) +
  geom_vline(xintercept = 0, color = "grey") +
  geom_linerange(aes(xmin = l, xmax = h), position = pos, linewidth = 2) +
  geom_linerange(aes(xmin = ll, xmax = hh), position = pos) +
  geom_point(position = pos, color = "black") +
  coord_cartesian(xlim = c(-.65, 1)) +
  bayesplot::theme_default() +
  ggtitle("Posterior distbributions (mean, CI66, CI95)")
```

### Effect sizes

Effect sizes of chosen model just to get exact values of means etc. if
necessary.

```{r effect-sizes, message = FALSE}
posterior1 %>%
  posterior::summarize_draws() %>%
  knitr::kable()
emm <- m_1 %>% emmeans(
  revpairwise ~ seeded_pool + herbicide | site,
  type = "response"
  )
emm$emmeans
```

```{r write-csv, echo = FALSE}
draws1 %>% write_csv(here("outputs", "statistics", "table_pool.csv"))
```

# Session info (BARG 2.A/6.A/6.B)

```{r session-info, echo = FALSE}
sessionInfo()
```
