---
title: "GREEEN tallgrass prairie restoration: RLQ analyses of pool size (Lux Arbor)"
author: "MB"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    dev: "png"
    fig_width: 7
    fig_height: 5
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
library(tinytex)
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation

#### Packages

```{r libraries, message = FALSE}
library(here)
library(tidyverse)
library(ade4)
```

```{r echo = FALSE}
rm(list = ls())
```

#### Load data

```{r load-data}
sites <- read_csv(
  here("data", "processed", "data_processed_sites.csv"),
  col_names = TRUE, na = c("na", "NA", ""), col_types = cols(
    .default = "?",
    id_plot = "f",
    year = "i",
    seeding_time = col_factor(
      levels = c("unseeded", "fall", "spring"), ordered = FALSE
      ),
    seeded_pool = col_factor(
      levels = c("0", "6", "12", "18", "33")#, ordered = TRUE
      ),
    treatment_id = "f",
    richness_type = "f"
  )
) %>%
  filter(
    site == "Lux Arbor",
    richness_type == "seeded_richness",
    !(treatment_id %in% c("2", "4"))
  ) %>%
    mutate(
    treatment = str_c(seeding_time, seeded_pool, sep = "_"),
    treatment = factor(treatment)
  ) %>%
  select(
    id_plot_year, year, treatment, water_cap, cover_non_seeded,
    cover_seeded_grass
    )

traits <- read_csv(
  here("data", "processed", "data_processed_traits.csv"),
  col_names = TRUE, na = c("na", "NA", ""), col_types = cols(.default = "?")
  ) %>%
  filter(seeded == "1", family != "Poaceae") %>%
  select(accepted_name, sla, height, seedmass) %>%
  mutate(seedmass = log(seedmass)) %>% 
  drop_na()

species <- read_csv(
  here("data", "processed", "data_processed_species.csv"),
  col_names = TRUE, na = c("na", "NA", ""), col_types = cols(.default = "?")
  ) %>%
  filter(plot_size %in% c("1", "25")) %>%
  mutate(abundance = 1) %>%
  select(id_plot_year, accepted_name, abundance) %>%
  semi_join(sites, by = "id_plot_year") %>%
  semi_join(traits, by = "accepted_name") %>%
  arrange(id_plot_year)
```

## Checks

#### Plots per year

```{r check-plot-number}
sites %>%
  group_by(year) %>%
  count()
```

#### Seeded species richness

```{r check-species-richness}
species %>%
  group_by(id_plot_year) %>%
  count(name = "seeded_forb_richness") %>%
  ungroup() %>% 
  count(seeded_forb_richness, name = "number_of_plots") %>%
  mutate(
    "plots_>=_1_species" = sum(number_of_plots),
    total_number_plots = 288
    )
```

#### Check non-occuring seeded species

```{r check-non-occuring-species}
traits %>% anti_join(species, by = "accepted_name")
```

## Finalize preparation

```{r finalize-preaparation}
traits <- traits %>%
  semi_join(species, by = "accepted_name") %>%
  column_to_rownames(var = "accepted_name")

sites <- sites %>%
  semi_join(species, by = "id_plot_year") %>%
  column_to_rownames(var = "id_plot_year")

species <- species %>%
  pivot_wider(
    names_from = "accepted_name", values_from = "abundance", values_fill = 0
    ) %>%
  column_to_rownames(var = "id_plot_year")
```

# Statistics

## Descriptive analyses

```{r descriptive-statistics}
sites %>%
  group_by(year) %>%
  count()
```


## Ordinations

```{r ordinations-for-rlq, warning=FALSE}
dudi_l <- dudi.coa(species, scannf = FALSE)
dudi_r <- dudi.hillsmith(
  sites, 
  row.w = dudi_l$lw,
  scannf = FALSE
  )
dudi_q <- dudi.pca(
  traits,
  row.w = dudi_l$cw,
  scannf = FALSE
  )
```


## RLQ analyses

```{r calculate-rlq, collapse = TRUE}
m_rlq <- rlq(
  dudiR = dudi_r, 
  dudiL = dudi_l, 
  dudiQ = dudi_q,
  scannf = FALSE
  )
```

### Global test

Model 2: species-environment relationship

Model 4: species-traits relationship

```{r rlq-global-test}
randtest(m_rlq, nrepet = 999, modeltype = 6)
```

### Traits by environment crossed table

```{r rlq-table}
m_rlq$tab 
```

### RLQ Plots

```{r rlq-plots}
plot(m_rlq)
s.label(m_rlq$lR)
s.label(m_rlq$lQ)
s.arrow(m_rlq$l1)
s.arrow(m_rlq$c1)
```

## Fourth-corner analysis

```{r calculate-4th-corner}
m_4th <- fourthcorner(
  tabR = sites, 
  tabL = species, 
  tabQ = traits,
  modeltype = 6,
  p.adjust.method.G = "none", 
  p.adjust.method.D = "none", 
  nrepet = 49999
)
m_4th # non-adjusted p-values
```

```{r adjsut-p-4th-corner}
m_4th_adj <- p.adjust.4thcorner(
  m_4th,
  p.adjust.method.G = "fdr", 
  p.adjust.method.D = "fdr", 
  p.adjust.D = "global"
)
m_4th_adj
```

### Plot

```{r plot-4th-corner-matrix}
plot(m_4th_adj, alpha = 0.05, stat = "D2")
```

```{r plot-4th-corner-biplot}
plot(m_4th_adj,
  x.rlq = m_rlq,
  alpha = 0.05,
  stat = "D2",
  type = "biplot"
)
```


# Save

```{r save, include = FALSE}
scores_sites <- m_rlq$lR %>%
  mutate(type = "sites") %>%
  rename(axis1 = 1, axis2 = 2)
scores_species <- m_rlq$lQ %>%
  mutate(type = "species") %>%
  rename(axis1 = 1, axis2 = 2)
scores_environment <- m_rlq$l1 %>%
  mutate(type = "environment") %>%
  rename(axis1 = 1, axis2 = 2)
scores_traits <- m_rlq$c1 %>%
  mutate(type = "traits") %>%
  rename(axis1 = 1, axis2 = 2)

scores <- bind_rows(
  scores_sites, scores_species, scores_environment, scores_traits
  ) %>%
  rownames_to_column(var = "id")

write_csv(
  scores, here(
    "outputs", "models", "model_rlq_pool_lux_arbor.csv"
    )
)
```


# Session info

```{r session-info, echo = FALSE}
sessionInfo()
```
